Stability and Performance of Coalitions of Prosumers Through Diversification in the Smart Grid

### 2016-02-05 Article.tex
respond letter

In the general case where distributions and correlation structures have no special form, finding an analytical expression for $ \alpha $ appears complicated. To overcome this problem, we seek an approximation for $ \alpha $ in a simplified situation where all power distributions are approximated by normal distributions and all prosumers are considered equivalent to a mean agent : $ \forall i \in \mathcal{A},\ P_i \sim \mathcal{N}(\bar{\mu}, \bar{\sigma}) $, where $ \bar{\mu} = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i $, $ \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i $, and $\forall i,j,\ \rho_{ij} = \bar{\rho} $. In this simplified case, we can express the utility as a function of $ \bar{\mu} $, $ \bar{\sigma} $, $\bar{\rho} $, $ \bar{N} $, and $ \alpha $ (see Appendix for more details). We then select $ \alpha^{\star} $ such that :

@@

Which leads to the following expression for $ \alpha^{\star} $ (see Appendix) :

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=.46]{./figs/figure_3}
		\caption{{\footnotesize Histograms of utility values for coalitions of size 3 ($\bar{N}=3$) in a $N=200$ prosumers example ($\phi = 0.1$, $ \alpha^{\star} = 0.08 $, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $). Red bars stand for cliques in the decorrelation graph, and blue bars for all the other possible triples. Cliques tend to exhibit higher utilities than random coalitions.}}
		\label{fig:histo_cliques}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=.46]{./figs/figure_4}
		\caption{{\footnotesize Utility of random coalitions depending on their size in a $N=200$ prosumers example ($\phi = .1$, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $). Blue dots show real mean utility values and the thick red curve its smoothed version by applying a Savitzky-Golay filter. On this plot, $ \alpha^{\star} = 0.006 $, and was selected according to eq. \ref{eq:alpha_star} in order to favor $\bar{N} = 40$ agents coalitions.}}
		\label{fig:real_utility2}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=.46]{./figs/figure_5}
		\caption{{\footnotesize Coalitions formed in the (contract value, volatility) space. The color map indicates the underlying quality. The closer to red the better (high contract values with small volatility). On the opposite, blue areas show poor quality (small contract values with high volatility). Blue dots stand for the decorrelated coalitions that we formed while green squares show correlated coalitions. The smaller yellow markers stand for the gravity centers of the coalition structures. The black dotted line shows how contract values and volatility evolve  when the size of the coalitions increases (a subset of the points are labeled by the size of the coalition they represent). Each point is the average over $ 10^{5} $ unconstrained draws of a random coalition. We also show the distribution of the gravity centers of random coalition structures with red ellipses (center is the mean, and each ellipse corresponds to one  standard deviation).}}
		\label{fig:coalitions}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=.46]{./figs/figure_6}
		\caption{{\footnotesize Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot)} }
		\label{fig:resilience}
	\end{center}
\end{figure}

@@

In this paper we studied how aggregations of prosumers could be authorized to sell their surplus of production to the grid operator. By relying on the past values of the agents, we constrained the market entry to both sufficiently productive and stable coalitions. The power that a coalition is able to propose on the market is therefore related to production and stability. The correlations between the prosumers impact directly the volatility of the coalitions, which led us to seek uncorrelated aggregations of agents. We used a graph representation of the correlation relationships between the agents as a reduced landscape for the coalition formation. A greedy algorithm that starts with cliques of the "de-correlation" graph of the agents and makes local improvements offers a good compromise between speed and quality of the results. We compared these results with random samplings, and an opposite strategy that clusters correlated agents together. We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility. Because of their better production over volatility ratios, these coalitions will tend to use less storage and waste less energy than more unstable coalitions. Because in real situations, agents are prone to failure, resilience is also an important criterion for the quality of the aggregations. We therefore studied how the coalitions remain on the market when their agents fail randomly. We showed that, in this situation, the coalitions resulting from our algorithm better withstand losses of agents.

>>>>

In the general case where distributions and correlation structures have no special form, finding an analytical expression for $ \alpha $ appears complicated. To overcome this problem, we seek an approximation for $ \alpha $ in a simplified situation where all power distributions are approximated by normal distributions and all prosumers are considered equivalent to a mean agent : $ \forall i \in \mathcal{A},\ P_i \sim \mathcal{N}(\bar{\mu}, \bar{\sigma}) $, where $ \bar{\mu} = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i $, $ \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i $, and $\forall i,j,\ \rho_{ij} = \bar{\rho} $. In this simplified case, we can express the utility as a function of $ \bar{\mu} $, $ \bar{\sigma} $, $\bar{\rho} $, $ \bar{N} $, and $ \alpha $ (see Appendix). We then select $ \alpha^{\star} $ such that :

@@

Which leads to (see Appendix) :

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=.45]{./figs/figure_3}
		\caption{{\footnotesize Histograms of utility values for coalitions of size 3 ($\bar{N}=3$) in a $N=200$ prosumers example ($\phi = 0.1$, $ \alpha^{\star} = 0.08 $, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $). Red bars stand for cliques in the decorrelation graph, and blue bars for all the other possible triples. Cliques tend to exhibit higher utilities than random coalitions.}}
		\label{fig:histo_cliques}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=.45]{./figs/figure_4}
		\caption{{\footnotesize Utility of random coalitions depending on their size in a $N=200$ prosumers example ($\phi = .1$, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $). Blue dots show real mean utility values and the thick red curve its smoothed version by applying a Savitzky-Golay filter. On this plot, $ \alpha^{\star} = 0.006 $, and was selected according to eq. \ref{eq:alpha_star} in order to favor $\bar{N} = 40$ agents coalitions.}}
		\label{fig:real_utility2}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=.45]{./figs/figure_5}
		\caption{{\footnotesize Coalitions formed in the (contract value, volatility) space. The color map indicates the underlying quality. The closer to red the better (high contract values with small volatility). On the opposite, blue areas show poor quality (small contract values with high volatility). Blue dots stand for the decorrelated coalitions that we formed while green squares show correlated coalitions. The smaller yellow markers stand for the gravity centers of the coalition structures. The black dotted line shows how contract values and volatility evolve  when the size of the coalitions increases (a subset of the points are labeled by the size of the coalition they represent). Each point is the average over $ 10^{5} $ unconstrained draws of a random coalition. We also show the distribution of the gravity centers of random coalition structures with red ellipses (center is the mean, and each ellipse corresponds to one  standard deviation).}}
		\label{fig:coalitions}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[scale=.45]{./figs/figure_6}
		\caption{{\footnotesize Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot)} }
		\label{fig:resilience}
	\end{center}
\end{figure}

@@

In this paper we studied how aggregations of prosumers could be authorized to sell their surplus of production to the grid operator. By relying on the past values of the agents, we constrained the market entry to both sufficiently productive and stable coalitions. The power that a coalition is able to propose on the market is therefore related to production and stability. The correlations between the prosumers impact directly the volatility of the coalitions, which led us to seek uncorrelated aggregations of agents. We used a graph representation of the correlation relationships between the agents as a reduced landscape for the coalition formation. A greedy algorithm that starts with cliques of the "de-correlation" graph of the agents and makes local improvements offers a good compromise between speed and quality of the results. We compared these results with random samplings, and an opposite strategy that clusters correlated agents together. We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility. Because of their better production over volatility ratios, these coalitions will tend to use less storage and waste less energy than more unstable coalitions. Because in real situations, agents are prone to failure, resilience is also an important criterion for the quality of the aggregations. We therefore studied how the coalitions remain on the market when their agents fail randomly. We showed that, in this situation, the coalitions resulting from our algorithm better withstand losses of agents. A possible direction for future work could be to use, in addition to real weather data, data from energy markets and aggregators as to propose a prosumer aggregation model closer to real conditions.
